\documentclass[11pt, a4paper]{article}


\usepackage{../mysty}

\renewcommand{\lesson}{0}
\renewcommand{\lessontitle}{Homework 2}
\renewcommand{\fulltitle}{\lessontitle}

\renewcommand{\themycount}{1.\arabic{mycount}}

\usepackage{xr}
\externaldocument{../lesson-03/2021b-toc-note-03}
\externaldocument{../lesson-05/2021b-toc-note-05}

%\usetikzlibrary{automata,positioning}
%\tikzset{initial text={}}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% START DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\date{}



\begin{center}
{\Large {\bf \fulltitle}}
\end{center}
\begin{center}
{\Large {\bf Due on Monday, 10:30 am, 25 April 2022 (111/04/25)}}
\end{center}


\vspace{0.5cm}

\paragraph*{Question 1 (1 point).}
Consider the problem $\circuiteval$ and Theorem~\ref{theo:circuit-eval} in Note~3.
The main idea for the proof of hardness is as follows.
Let $\cM$ be a polynomial time DTM.
On input word $w$, the reduction constructs a circuit $C$ such that:
\begin{eqnarray*}
C(w)=1 \qquad& \text{if and only if} &\qquad \cM\ \text{accepts}\ w
\end{eqnarray*} 
The construction of $C$ is similar to Cook-Levin reduction for the proof of $\npt$-hardness of $\sat$.
Explain why we need the fact that $\cM$ is DTM and {\em not} NTM in the proof of Theorem~\ref{theo:circuit-eval}.




\paragraph*{Question 2 (2 points).}
Prove that $\pht$ collapses if any one of the following is true.
\begin{itemize}
\item
$\sigmap {i} = \pip {i}$ for some $i\geq 1$.
\item
There is $\pht$-complete language.
\item
$\pht=\ps$.
\end{itemize}

\paragraph*{Question 3 (2 points).}
Suppose that $A$ is a language such that $\pt^A =\npt^A$.
Prove that $\pht^{A}\subseteq \pt^A$.

\vspace{0.3cm}
{\bf Def:} A language $L$ is in the class {\em $\sigmap {i}$ with oracle access to $A$},
if there is a polynomial $q(n)$ and a polynomial time DTM $\cM^A$
such that for every $w\in \Sigma^*$ the following holds.
\begin{eqnarray*}
\label{eq:sigmap}
w \in L & \text{iff} &
\exists y_1 \in \{0,1\}^{q(|w|)}\ 
\forall y_2 \in \{0,1\}^{q(|w|)}\ 
\cdots \
Q y_i \in \{0,1\}^{q(|w|)}\
\cM^A \ \text{accepts}\ (w,y_1,\ldots,y_i) 
\end{eqnarray*}
As before, $\cM^{A}$ denotes the TM $\cM$ with oracle access to $A$.
A language $L$ is in $\pht^A$, if there is $i$ such that $L$ is in the class $\sigmap {i}$ with oracle access to $A$.


\paragraph*{Question 4 (1 point).}
Consider the following two problems.
\begin{quote}
{\def\arraystretch{1.25}
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|l|}{{\sf CYCLE-COVER}}
\\
\hline
{\bf Input:}
&
A directed graph $G$.
\\
{\bf Task:}
&
Output a cycle cover of $A$, if exists.
Otherwise, output $0$.
\\
\hline
\end{tabular}}
\end{quote}
\begin{quote}
{\def\arraystretch{1.25}
\begin{tabular}{|ll|}
\hline
\multicolumn{2}{|l|}{{\sf MATCHING}}
\\
\hline
{\bf Input:}
&
A bipartite (undirected) graph $H=(U,V,E)$, where $|U|=|V|$.
\\
{\bf Task:}
&
Output a matching of $H$, if exists.
Otherwise, output $0$.
\\
\hline
\end{tabular}}
\end{quote}
Here a matching of $H=(U,V,E)$ is defined as a set $E_0\subseteq E$
such that every vertex $u \in U\cup V$ is incident to exactly one edge in $E_0$.
In other words, $E_0$ forms a ``bijection'' from $U$ to $V$.

Prove that {\sf CYCLE-COVER} and {\sf MATCHING} are reducible to each other in polynomial time.








\paragraph*{Question 5 (2 points).}
Prove Lemma~\ref{lem:xor-graph} in Note 5.



\paragraph*{Question 6 (2 points).}
Prove Lemma~\ref{lem:cycle-cover-assignment} in Note 5.



\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%% END OF DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%





